---
format:
  gfm:
    default-image-extension: "svg"
---

<!-- README.md is generated from README.qmd. Edit README.qmd. -->

```{r}
#| include: false
knitr::opts_chunk$set(
  collapse = TRUE,
  comment = "#>",
  fig.path = "man/figures/README-",
  out.width = "100%",
  fig.ext = "svg",
  dev = "svglite"
)
```

# delphiPalettes

<!-- badges: start -->
[![R-CMD-check](https://github.com/ConnorB/delphiPalettes/actions/workflows/R-CMD-check.yaml/badge.svg)](https://github.com/ConnorB/delphiPalettes/actions/workflows/R-CMD-check.yaml)
[![Codecov test coverage](https://codecov.io/gh/ConnorB/delphiPalettes/graph/badge.svg)](https://app.codecov.io/gh/ConnorB/delphiPalettes)
<!-- badges: end -->

delphiPalettes provides `r length(delphiPalettes::delphi_palettes())` color palettes from the [delphi.tools](https://delphi.tools) [palette collection](https://delphi.tools/tools/palette-collection).

## Installation

Install the development version from [GitHub](https://github.com/ConnorB/delphiPalettes).

Load the package.

``` r
# install.packages("pak")
pak::pak("ConnorB/delphiPalettes")
```
```{r}
#| label: setup
library(delphiPalettes)
```

## Browse palettes

`print_delphi_palettes()` draws one panel for each palette. Set `category` or `colorblind_friendly = TRUE` to limit the result.

<!-- The raw HTML `alt` attribute supplies the alt text. The Quarto GFM
     renderer does not connect `fig-alt` to this attribute:
     https://github.com/quarto-dev/quarto-cli/issues/12456 -->

```{r}
#| label: quick-look
#| out.extra: >-
#|   alt="Palette grid with 34 panels in six rows and six columns. Each
#|   panel has a keycaps palette name. For example: &quot;2600&quot;,
#|   &quot;dualshot&quot;, &quot;nautilus&quot;, and &quot;vaporwave&quot;. Each panel has
#|   a horizontal strip of palette colors."

print_delphi_palettes("keycaps")
```

The `keycaps` category has `r length(delphiPalettes::delphi_palettes("keycaps"))` palettes. The collection has `r length(delphiPalettes::delphi_palettes())` palettes in ten categories. With no filters, `print_delphi_palettes()` draws one page per category.

## Get palette colors

Each palette name uses lowercase letters and hyphens. For example, use `"mayfair"` or `"grand-budapest"`. `delphi_palette()` returns a character vector of hexadecimal colors.

```{r}
delphi_palette("mayfair")
```

`print_delphi_palette()` draws the palette.

```{r}
#| label: print-example
#| out.extra: >-
#|   alt="Palette strip with five color blocks. The label is
#|   &quot;mayfair&quot;. The colors progress from dark navy and green to
#|   maroon, tan, and cream."

print_delphi_palette("mayfair")
```

Set `direction = -1` to reverse the color order. Set `n` to return the first colors in the selected order.

```{r}
delphi_palette("grand-budapest", direction = -1)
delphi_palette("pride", n = 3)
```

Set `type = "continuous"` to interpolate a palette. You must also set `n`.

```{r}
continuous <- delphi_palette("mayfair", n = 20, type = "continuous")
continuous
```

```{r}
#| label: continuous-example
#| echo: false
#| out.extra: >-
#|   alt="Palette strip with 20 adjacent color blocks from the
#|   &quot;mayfair&quot; palette. The colors progress from dark navy and green
#|   to maroon, tan, and cream."

op <- par(mar = c(0, 0, 0, 0))
image(
  seq_along(continuous),
  1,
  as.matrix(seq_along(continuous)),
  col = continuous,
  xaxt = "n",
  yaxt = "n",
  xlab = "",
  ylab = "",
  bty = "n"
)
par(op)
```

### Use with ggplot2

`scale_color_delphi_d()` and `scale_fill_delphi_d()` map factor levels to palette colors. If a factor has more levels than colors, these scales interpolate colors. `scale_color_delphi_c()` and `scale_fill_delphi_c()` map numeric values to palette colors. `scale_colour_delphi_d()` and `scale_colour_delphi_c()` use British spelling.

```{r}
#| label: ggplot-discrete
#| message: false
#| out.extra: >-
#|   alt="Box plot. The x-axis has iris species. The y-axis has sepal
#|   length. The boxes use red, orange, and yellow from the pride palette.
#|   Median sepal length increases from setosa to virginica."
library(ggplot2)

ggplot(iris, aes(Species, Sepal.Length, fill = Species)) +
  geom_boxplot() +
  scale_fill_delphi_d("pride") +
  theme_minimal() +
  theme(legend.position = "none")
```

```{r}
#| label: ggplot-continuous
#| out.extra: >-
#|   alt="Scatter chart. The x-axis has sepal length. The y-axis has
#|   sepal width. Color represents petal length from 1 to 6.9 on the
#|   mayfair scale. Petal length increases as sepal length increases."

ggplot(iris, aes(Sepal.Length, Sepal.Width, color = Petal.Length)) +
  geom_point(size = 2) +
  scale_color_delphi_c("mayfair") +
  theme_minimal()
```

### Categories

The collection has ten categories: `classic`, `nature`, `keycaps`, `vintage`, `modern`, `bold`, `soft`, `monochrome`, `seasonal`, and `artistic`. Use `delphi_palettes()` to get one category.

```{r}
names(delphi_palettes("keycaps"))
```

### Color vision support

`{colorblindcheck}` simulates deuteranopia, protanopia, and tritanopia. `delphi_palette_colorblind()` reports the stored result for one palette. `delphi_palettes()` returns the `r length(delphiPalettes::delphi_palettes(colorblind_friendly = TRUE))` palettes that pass all three simulations.

```{r}
delphi_palette_colorblind("dove")
delphi_palette_colorblind("pinata")
```

```{r}
#| label: colorblind-grid
#| out.extra: >-
#|   alt="Palette grid with 21 panels in five rows and five columns.
#|   The palettes pass the three color vision simulations. Each panel
#|   has a palette name and a horizontal strip of palette colors."

print_delphi_palettes(colorblind_friendly = TRUE)
```

## Credits

[delphi.tools](https://delphi.tools) created the palette collection. Ruby Morgan Voigt created the palette designs. delphiPalettes makes these palettes available in R. [delphitools](https://github.com/1612elphi/delphitools) uses the MIT license. [LICENSE.md](LICENSE.md) includes its license notice.
