---
format:
  gfm:
    default-image-extension: "svg"
---

<!-- README.md is generated from README.qmd. Please edit that file -->

```{r}
#| include: false
knitr::opts_chunk$set(
  collapse = TRUE,
  comment = "#>",
  fig.path = "man/figures/README-",
  out.width = "100%",
  fig.ext = "svg",
  dev = "svglite"
)
```

# delphiPalettes

<!-- badges: start -->
[![R-CMD-check](https://github.com/ConnorB/delphiPalettes/actions/workflows/R-CMD-check.yaml/badge.svg)](https://github.com/ConnorB/delphiPalettes/actions/workflows/R-CMD-check.yaml)
[![Codecov test coverage](https://codecov.io/gh/ConnorB/delphiPalettes/graph/badge.svg)](https://app.codecov.io/gh/ConnorB/delphiPalettes)
<!-- badges: end -->

A collection of `r length(delphiPalettes::delphi_palettes())` color palettes for data visualisation, ported from the [palette collection](https://delphi.tools/tools/palette-collection) at [delphi.tools](https://delphi.tools) — a suite of privacy-respecting, browser-based design tools created by [delphi](https://github.com/1612elphi/delphitools). All palette designs are delphi's; this package just makes them usable from R.

## Installation

You can install the development version of delphiPalettes from [GitHub](https://github.com/) with:

``` r
# install.packages("pak")
pak::pak("ConnorB/delphiPalettes")
```
```{r}
#| label: setup
library(delphiPalettes)
```

## A quick look

`print_delphi_palettes()` draws a grid of palette swatches, one panel per palette. Pass a `category` or `colorblind_friendly = TRUE` to see a manageable slice of the collection:

<!-- alt text is set via out.extra (a raw HTML `alt` attribute) rather than
     fig-alt, which quarto's gfm renderer doesn't wire up to `alt`:
     https://github.com/quarto-dev/quarto-cli/issues/12456 -->

```{r}
#| label: quick-look
#| out.extra: >-
#|   alt="Grid of 34 color palette swatches, one panel per palette in
#|   the keycaps category, arranged 6 panels by 6 panels. Each panel
#|   is labeled with its palette name, such as &quot;2600&quot;,
#|   &quot;dualshot&quot;, &quot;nautilus&quot;, and &quot;vaporwave&quot;,
#|   above a horizontal strip of solid color blocks."

print_delphi_palettes("keycaps")
```

That's the `keycaps` category (`r length(delphiPalettes::delphi_palettes("keycaps"))` palettes); there are `r length(delphiPalettes::delphi_palettes())` in total across ten categories. Leave `category` unset to grid the whole collection at once (a large plot).

## Usage

Every palette has a lowercase, hyphenated name, like `"mayfair"` or `"grand-budapest"`. `delphi_palette()` returns one as a character vector of hex colors:

```{r}
delphi_palette("mayfair")
```

`print_delphi_palette()` draws it:

```{r}
#| label: print-example
#| out.extra: >-
#|   alt="Horizontal strip of 5 solid color blocks labeled
#|   &quot;mayfair&quot; above the strip. The colors move from dark
#|   navy blue and forest green through a deep maroon red to a warm
#|   tan and pale cream."

print_delphi_palette("mayfair")
```

Reverse the color order with `direction = -1`, or take a subset with `n`:

```{r}
delphi_palette("grand-budapest", direction = -1)
delphi_palette("pride", n = 3)
```

Interpolate a palette to any number of colors with `type = "continuous"`:

```{r}
continuous <- delphi_palette("mayfair", n = 20, type = "continuous")
continuous
```

```{r}
#| label: continuous-example
#| echo: false
#| out.extra: >-
#|   alt="Horizontal gradient bar showing the mayfair palette
#|   interpolated to 20 continuous color steps. The gradient blends
#|   smoothly from dark navy blue through forest green and deep
#|   maroon red to tan and pale cream, with no visible banding
#|   between adjacent steps."

op <- par(mar = c(0, 0, 0, 0))
image(
  seq_along(continuous),
  1,
  as.matrix(seq_along(continuous)),
  col = continuous,
  xaxt = "n",
  yaxt = "n",
  xlab = "",
  ylab = "",
  bty = "n"
)
par(op)
```

### Use with ggplot2

Discrete scales (`scale_color_delphi_d()` / `scale_fill_delphi_d()`) extend a palette by interpolation if there are more levels than colors; continuous scales (`scale_color_delphi_c()` / `scale_fill_delphi_c()`) interpolate smoothly. `scale_colour_*()` spellings are provided as aliases.

```{r}
#| label: ggplot-discrete
#| message: false
#| out.extra: >-
#|   alt="Box plot chart. Iris species (setosa, versicolor, virginica)
#|   along the x-axis, sepal length along the y-axis. Each box is
#|   filled with a different color from the pride palette — red,
#|   orange, and yellow. Median sepal length rises from setosa through
#|   versicolor to virginica, and the spread of values widens across
#|   the same order."
library(ggplot2)

ggplot(iris, aes(Species, Sepal.Length, fill = Species)) +
  geom_boxplot() +
  scale_fill_delphi_d("pride") +
  theme_minimal() +
  theme(legend.position = "none")
```

```{r}
#| label: ggplot-continuous
#| out.extra: >-
#|   alt="Scatter chart. Sepal length along the x-axis, sepal width
#|   along the y-axis, points colored on a continuous mayfair scale by
#|   petal length (ranging from 1 to 6.9). Points cluster loosely with
#|   no strong linear trend between sepal length and width, but color
#|   shifts from the dark end of the scale at low sepal length values
#|   to the pale end at high sepal length values, reflecting the
#|   strong correlation between petal length and sepal length."

ggplot(iris, aes(Sepal.Length, Sepal.Width, color = Petal.Length)) +
  geom_point(size = 2) +
  scale_color_delphi_c("mayfair") +
  theme_minimal()
```

### Categories

Palettes are grouped into ten categories — `classic`, `nature`, `keycaps`, `vintage`, `modern`, `bold`, `soft`, `monochrome`, `seasonal`, and `artistic`. Pass one to `delphi_palettes()` to see just that group:

```{r}
names(delphi_palettes("keycaps"))
```

### Colorblind-friendly palettes

Every palette has been checked with [colorblindcheck](https://cran.r-project.org/package=colorblindcheck) for whether it stays distinguishable under simulated deuteranopia, protanopia, and tritanopia. `delphi_palette_colorblind()` reports the verdict for one palette, and `delphi_palettes(colorblind_friendly = TRUE)` filters to only the `r length(delphiPalettes::delphi_palettes(colorblind_friendly = TRUE))` palettes that pass all three:

```{r}
delphi_palette_colorblind("dove")
delphi_palette_colorblind("pinata")
```

```{r}
#| label: colorblind-grid
#| out.extra: >-
#|   alt="Grid of 21 color palette swatches, one panel per palette
#|   that passes deuteranopia, protanopia, and tritanopia checks,
#|   arranged 5 panels by 5 panels with 4 empty panels at the end.
#|   Each panel is labeled with its palette name, such as
#|   &quot;hokitika&quot;, &quot;obsidian&quot;, &quot;porcelain&quot;,
#|   and &quot;escher&quot;, above a horizontal strip of solid color
#|   blocks."

print_delphi_palettes(colorblind_friendly = TRUE)
```

## Credit

The palette designs in this package are not original work — they're ported from the palette collection built into [delphi.tools](https://delphi.tools) ([source](https://github.com/1612elphi/delphitools)), created by delphi (Ruby Morgan Voigt). delphiPalettes just packages that collection for use in R (retrieval, interpolation, and colorblind-safety checks); all credit for the palettes themselves belongs to the original designer. delphitools is MIT licensed; its licence notice is reproduced alongside this package's own in [LICENSE.md](LICENSE.md).
